{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-10-06 thesis/future.tex
future heuristics qs/icb

There is much room to expand Landslide's use of heuristics to find bugs as soon as possible should they exist.
% TODO
incorporate ICB \cite{chess-icb}
% ICB incorporations
% into quicksand - when SS estimates are too high, fall back on a preemption bounded version to find bugs - improve the partial verif
% into education - provide more partial verifs to students (but is it clear they'll understand the subtleties of it?)
% into txn - incorporate failure injection into the model

% arugment why ML is NOT appropriate for smc in general -- if you try to train it to choose thread interleavings / search ordering,
% even if you're already in not-enough-cpu-time heuristics land, you're biasing it to find only the types of bugs you've already seen
% and destroying its ability to find novelties
% if your ML provides a partial verification, it is impossible for a human to understand what that verification means
As a matter of personal philosophy,
I would strongly caution against using machine learning to try to expose bugs faster.
Machine learning has become popular recently as a magic cure-all for many programming problems,
but is also notorious for amplifying any biases in its training set in often unpredictable ways
\cite{conceptnet-bias}.
In the domain of concurrency testing, this would translate to being able to find only types of bugs
already seen in the training data, abandoning any novel bug patterns to further obscurity.
Granted, Iterative Deepening and ICB also deprioritize ``deeper'' bugs,
as measured by number of unique or total preemptions, respectively,
but this strategy is at least backed up by empirical evidence
% TODO: upgrade citation to also have a section reference to that icb bound table from quicksand eval
\cite{chess-icb},
and when a bug is not found,
the strategy is transparent enough that the user can understand the nature of the partial verification.
%
Machine learning would be more appropriate for improving state space estimation,
whose output is already a black-box oracle from the user's perspective,
and where the worst case of being wrong
is that Quicksand prioritize its jobs suboptimally. % subjunctive
In-progress estimates fluctuate depending on many factors,
such as interleaving-dependent flow control,
and I suspect deeper patterns may exist among multiple different state spaces,
for example,
from the same test program with different $K,N$ parameters.
For any future researchers who wish to study such patterns,
% with machine learning... don't feel like i can do justice to discussing its flaws with real citations
I have made the estimation logs from the HTM experiments (\sect{\ref{sec:tm-eval}})
freely available at \url{https://github.com/bblum/thesis-htm-logs}.
%
Nevertheless,
any overall search heuristic for concurrency testing
must be designed to support the human user's agency first and foremost,
that she may readily understand partial verification results
and participate effectively in the test case refinement strategies discussed
upchapter. % this is a word now.

>>>>

There is much room to expand Landslide's use of heuristics to balance verification with fast bug-finding.
Landslide's maximal state space mode (\sect{\ref{sec:tm-verif}}) currently requires the user to decide in advance
whether she thinks a full verification is possible within the time limit,
and supply {\tt -M} if so,
sacrificing some bug-finding power up front
(or just running the test twice).
Quicksand could make this decision automatically,
always dedicating one of its available CPUs to the maximal state space
while the rest run Iterative Deepening as normal.
Currently, Landslide explores each state space sequentially,
but if extended with Distributed DPOR \cite{parallel-dpor},
Quicksand could also dynamically allocate more or fewer CPUs to the maximal state space according to its ETA.
% ICB incorporations
% into quicksand - when SS estimates are too high, fall back on a preemption bounded version to find bugs - improve the partial verif
% into education - provide more partial verifs to students (but is it clear they'll understand the subtleties of it?)
% into txn - incorporate failure injection into the model
Likewise, Quicksand could
incorporate ICB \cite{chess-icb}
to get the best of both worlds:
when testing smaller data-race jobs,
start them in ICB mode to begin with,
and when the ETAs of larger jobs (including the maximal one) grow too large,
downgrade them to ICB to try to at least get a partial verification for that preemption point set
rather than discarding it entirely.
ICB could also be extended to include HTM failure injections as part of its model,
counting them either as part of the overall preemption bound
or under a separate abort bound.
State space estimation under ICB currently
counts only DPOR's tagged branches that don't get skipped under the preemption bound,
estimating that bound's state space size
and resetting whenever the bound increases;
it could also be extended to be ICB-aware,
counting branches skipped due to the bound separately
and computing two estimates at once,
which would in turn allow Quicksand to make more informed decisions about when to use ICB.

% arugment why ML is NOT appropriate for smc in general -- if you try to train it to choose thread interleavings / search ordering,
% even if you're already in not-enough-cpu-time heuristics land, you're biasing it to find only the types of bugs you've already seen
% and destroying its ability to find novelties
% if your ML provides a partial verification, it is impossible for a human to understand what that verification means
As a matter of personal philosophy,
I would strongly caution against using machine learning to try to expose bugs faster.
Machine learning has become popular recently as a magic cure-all for many programming problems,
but is also notorious for amplifying any biases in its training set in often unpredictable ways
\cite{conceptnet-bias}.
In the domain of concurrency testing, this would translate to being able to find only types of bugs
already seen in the training data, abandoning any novel bug patterns to further obscurity.
Granted,
Iterative Deepening and
ICB also deprioritize ``deeper'' bugs,
as measured by number of unique or total preemptions, respectively,
%in terms of minimum necessary preemption count,
%but this strategy is informed by empirical evidence
%% TODO: upgrade citation to also have a section reference to that icb bound table from quicksand eval
%\cite{chess-icb},
%and
but when a bug is not found
these strategies are also transparent enough that the user can understand the nature of the partial verification.
%
Machine learning would be more appropriate for improving state space estimation,
whose output is already a black-box oracle from the user's perspective,
and where the worst case of being wrong
is that Quicksand prioritize its jobs suboptimally. % subjunctive
In-progress estimates fluctuate depending on many factors,
such as interleaving-dependent flow control,
and I suspect deeper patterns may exist among multiple different state spaces,
for example,
from the same test program with different $K,N$ parameters.
For any future researchers who wish to study such patterns,
% with machine learning... don't feel like i can do justice to discussing its flaws with real citations
I have made the estimation logs from the HTM experiments (\sect{\ref{sec:tm-eval}})
freely available at \url{https://github.com/bblum/thesis-htm-logs}.
%
Nevertheless,
any overall search heuristic for concurrency testing
must be designed to support the human user's agency first and foremost,
that she may readily understand partial verification results
and participate effectively in the test case refinement strategies discussed
upchapter. % this is a word now.
